Statically Typed String Sanitation Inside a Python

### 2014-08-14 sanitation-psp14/sanitation-psp14-short.tex

\usepackage{todonotes}

\usepackage{placeins}

@@

% Judgements
\newcommand{\err}{\ {\sf err}}
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

\begin{figure}[t]\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash s:\sigma}

\inferrule[S-T-Abs]
{\sctx, x : \sigma_1 \vdash e : \sigma_2}
{\sctx \vdash \lambda x:\sigma_1.e : \sigma_1 \rightarrow \sigma_2}

@@

\addtocounter{figure}{4}
\begin{figure*}[t]\label{fig:trans}

$\fbox{\inferrule{}{\trden{S} = P}}$
\begin{mathpar}

@@

\inferrule[Tr-SafeCoerce]
{ \ }
{ \sctx \vdash  \trden{ \coerce{r'}{e} } = \trden{e} }

\inferrule[Tr-Check-Ok]
{ \sctx \vdash e : \stringin{r} \\ \lang{r} \subseteq \lang{r'} } 
{ \sctx \vdash  \trden{ \rcheck{r'}{e}{x.e_1}{e_2} } = \trden{[e/x]e_1} }

\inferrule[Tr-Check-NotOk]
{ \sctx \vdash e : \stringin{r} \\ \lang{r} \not \subseteq \lang{r'}}
{ \sctx \vdash  \trden{ \rcheck{r'}{e}{x.e_1}{e_2} } = \trden{\tcheck{r'}{e_2}} }

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed in the Tr-SafeCoerce case.}
\end{figure*}
\addtocounter{figure}{-5}

@@

\inferrule[P-T-Abs]
{\tctx, x : \iota_1 \vdash e : \iota_2}
{\tctx \vdash \lambda x:\iota_1.e : \iota_1 \rightarrow \iota_2}

\inferrule[P-T-App]
{\tctx \vdash e_1 : \iota_2 \rightarrow \iota \\ \tctx \vdash e_2 : \iota_2}
{\tctx \vdash e_1(e_2) : \iota}

@@

\begin{figure}[t]
\small
$\fbox{\inferrule{}{P \treduces P}}$
$\fbox{\inferrule{}{P \err}}$

@@

\inferrule[P-E-App]
{ e_1 \sreduces e_1' \\  e_2 \sreduces e_2' }
{ e_1 e_2 \sreduces e_1' e_2'}

@@

\inferrule[P-E-Check-Err]
{\iota_1 \treduces \rx{r} \\ \iota_2 \treduces \tstr{s} \\ s \not \in \lang{r}}
{\tcheck{\iota_1}{\iota_2} \err}
\end{mathpar}
\caption{Big step semantics for of $\lambdap$.
Error propagation rules are omitted.}
\label{fig:ieval}
\end{figure}

@@

def main(db):
  input = sanitize(user_input())
  results = db.execute(results_query(input))
  return results_div(input) + format(results)
\end{lstlisting}
\vspace{-10px}
\caption{Regular string types in atlang, a library that enables static type checking for Python. }\label{fig:atexample}
\end{figure}
\begin{figure}
\begin{lstlisting}
class string_in(atlang.Type):
  def __init__(self, rx):
    rx = rx_normalize(rx)
    atlang.Type.__init__(idx=rx)

@@

\subsection{An Example}

Figure \ref{fig:atexample} demosntrates the use of string types in atlang.
The sanitize function takes an arbitrary string and returns a string without
double quotes or left and right brackets. In this example, we use HTML
escape sequences.

>>>>

%\usepackage{todo}
%\usepackage[subject={Todo},author={Josef}]{pdfcomment}
%\usepackage{cooltooltips}
\newcommand{\todo}[1]{{\color{red} #1}}
\usepackage{placeins}

@@

% Judgements
\newcommand{\trden}[1]{\llbracket #1 \rrbracket} % = Translation Denotation.

@@

\begin{figure}[t]\label{fig:lambdas}
\small
$\fbox{\inferrule{}{\sctx \vdash e : \sigma}}$
~~~~$\sctx ::= \emptyset \pipe \sctx, x : \sigma$
\begin{mathpar}
\inferrule[S-T-Var]
{ x:\sigma \in \sctx }
{ \sctx \vdash x:\sigma}

\inferrule[S-T-Abs]
{\sctx, x : \sigma_1 \vdash e : \sigma_2}
{\sctx \vdash \lambda x.e : \sigma_1 \rightarrow \sigma_2}

@@

\begin{figure*}[t]\label{fig:trans}

$\fbox{\inferrule{}{\sctx \vdash \trden{S} = P}}$
\begin{mathpar}

@@

\inferrule[Tr-SafeCoerce]
{ \sctx \vdash \trden{e} = \iota }
{ \sctx \vdash  \trden{ \coerce{r'}{e} } = \iota }

\inferrule[Tr-Check-Ok]
{ \sctx \vdash e : \stringin{r} \\ \lang{r} \subseteq \lang{r'} \\ \trden{[e/x]e_1} = \iota } 
{ \sctx \vdash  \trden{ \rcheck{r'}{e}{x.e_1}{e_2} } = \iota }

\inferrule[Tr-Check-NotOk]
{ \sctx \vdash e : \stringin{r} \\ \lang{r} \not \subseteq \lang{r'} \\ \trden{\tcheck{r'}{e_2}} = \iota}
{ \sctx \vdash  \trden{ \rcheck{r'}{e}{x.e_1}{e_2} } = \iota}

\end{mathpar}
\caption{Translation from source terms ($e$) to target terms ($\iota$).
The translation is type-directed.}
\end{figure*}

@@

\inferrule[P-T-Abs]
{\tctx, x : \tau_1 \vdash \iota_2 : \tau_2}
{\tctx \vdash \lambda x.\iota_2 : \tau_1 \rightarrow \tau_2}

\inferrule[P-T-App]
{\tctx \vdash \iota_1 : \tau_2 \rightarrow \tau \\ \tctx \vdash \iota_2 : \tau_2}
{\tctx \vdash \iota_1(\iota_2) : \iota}

@@

\begin{figure}[t]
\small
$\fbox{\inferrule{}{\iota \treduces \iota}}$

@@

\inferrule[P-E-App]
{ e_1 \sreduces e_1' \\  e_2 \sreduces e_2' }
{ e_1(e_2) \sreduces e_1'(e_2')}

@@

\end{mathpar}
\caption{Big step semantics for of $\lambdap$}.
\label{fig:ieval}
\end{figure}

@@

@fn
def main(db):
  input = sanitize(user_input())
  results = db.execute(results_query(input))
  return results_div(input) + format(results)
\end{lstlisting}
\vspace{-10px}
\caption{Regular string types in atlang, a library that enables static type checking for Python. }\label{fig:atexample}
\end{figure}
\begin{figure}
\begin{lstlisting}
class string_in(atlang.Type):
  def __init__(self, rx):
    rx = rx_normalize(rx)
    atlang.Type.__init__(idx=rx)

@@

\subsection{Example Usage}

Figure \ref{fig:atexample} demonstrates the use of two type constructors, \verb|fn| and \verb|string_in|, both of which we have included in \verb|atlib|, the standard library for \verb|atlang|. The \verb|fn| type constructor can be used to annotate functions that should be statically checked by \verb|atlang|. The function \verb|sanitize| on lines 3-7 specifies one argument, \verb|s|, of type \lstinline{string_in[r'.*']}.\todo{fix sizes}

The sanitize function takes an arbitrary string and returns a string without
double quotes or left and right brackets. In this example, we use HTML
escape sequences.
